Machine Learning Methods for Discourse Segmentation of Communications in E-Mail Based Behavioral Interventions

### 2018-07-24 paper-amia-summit-2018/manuscript.tex
Update manuscript

\noindent{\bf Abstract}
\textit{Communication science approaches to developing effective behavior interventions, such as motivational interviewing (MI), are limited by the traditionally manual qualitative coding of communication exchanges, which is a very resource-intensive and time-consuming process. This study focuses on the analysis of e-Coaching sessions, behavior interventions that are delivered via email and grounded in the principles of MI. A critical step towards automated annotation of e-Coaching communication exchanges is segmentation of emails into textual fragments that correspond to MI behaviors. In this work, we formulate this task as a classification problem and propose lexical, punctuation and semantic features to address it. We experimented both with traditional machine learning method, Conditional Random Field (CRF) and deep learning methods, such as Multilayer Perceptrons (MLPs), Bidirectional Recurrent Neural Networks (BRNNs) and Convolutional Recurrent Neural Networks (CRNNs). Results indicate that CRNN outperformed CRF, MLP and BRNN achieving 0.990 macro F1-score overall and 0.821 macro F1-score for detecting new segment.}

@@

We utilized three types of features in conjunction with CRF, MLP, BRNN and CRNN methods: lexical, punctuation and topic features. Lexical features represent each word in a pair as a binary vector, in which 1 corresponds to the word in question and 0 to all other words. Since part-of-speech (PoS) tags \cite{kotov2015interpretable,hashimoto2016topic,lu2016modeling} have been shown to be effective semantic abstractions of individual words, we used topic features for our experiment. Punctuation marks, which correspond to one of the symbols \{`.', `,', `!', `?', `:', `;', `-'\} between a pair of words, are also employed as a feature since punctuation marks designate the boundary of a sentence, clause and phrase and often also correspond to segment boundary.

@@

\textbf{Multilayer Perceptrons}: this method considers each training sample as a point in the input feature space. For a new test sample, Euclidean distance is calculated to find the k-nearest classified neighbors from the training set. Finally, the test sample is classified into a majority class of its k-nearest neighbors. We experimentally determined that best performance was achieved when the number of neighbors is 3.

>>>>

\noindent{\bf Abstract}
\textit{Communication science approaches to developing effective behavior interventions, such as motivational interviewing (MI), are limited by the traditionally manual qualitative coding of communication exchanges, which is a very resource-intensive and time-consuming process. This study focuses on the analysis of e-Coaching sessions, behavior interventions that are delivered via email and grounded in the principles of MI. A critical step towards automated annotation of e-Coaching communication exchanges is segmentation of emails into textual fragments that correspond to MI behaviors. In this work, we formulate this task as a classification problem and propose lexical, punctuation and part-of-speech features to address it. We experimented both with traditional machine learning method, Conditional Random Field (CRF) and deep learning methods, such as Multilayer Perceptrons (MLPs), Bidirectional Recurrent Neural Networks (BRNNs) and Convolutional Recurrent Neural Networks (CRNNs). Results indicate that CRNN outperformed CRF, MLP and BRNN achieving 0.990 macro F1-score overall and 0.821 macro F1-score for detecting new segment.}

@@

We utilized three types of features in conjunction with CRF, MLP, BRNN and CRNN methods: lexical, punctuation and topic features. Lexical features represent each word in a pair as a binary vector, in which 1 corresponds to the word in question and 0 to all other words. Since part-of-speech (PoS) tags \cite{kotov2015interpretable,hashimoto2016topic,lu2016modeling} have been shown to be effective semantic abstractions of individual words, we used topic features for our experiment. Punctuation marks, which correspond to one of the symbols \{`.', `,', `!', `?', `:', `;'\} between a pair of words, are also employed as a feature since punctuation marks designate the boundary of a sentence, clause and phrase and often also correspond to segment boundary.

@@

\textbf{Deep Learning Methods}: this method considers each training sample as a point in the input feature space. For a new test sample, Euclidean distance is calculated to find the k-nearest classified neighbors from the training set. Finally, the test sample is classified into a majority class of its k-nearest neighbors. We experimentally determined that best performance was achieved when the number of neighbors is 3.
